### 2016-04-11 Paper/Infovis Paper.tex

\documentclass{sigchi}
% Use this command to override the default ACM copyright statement
% (e.g. for preprints).  Consult the conference website for the
% camera-ready copyright statement.

%% HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP --
%% Please note you need to make sure the copy for your specific
%% license is used here!
% \toappear{
% Permission to make digital or hard copies of all or part of this work
% for personal or classroom use is granted without fee provided that
% copies are not made or distributed for profit or commercial advantage
% and that copies bear this notice and the full citation on the first
% page. Copyrights for components of this work owned by others than ACM
% must be honored. Abstracting with credit is permitted. To copy
% otherwise, or republish, to post on servers or to redistribute to
% lists, requires prior specific permission and/or a fee. Request
% permissions from \href{mailto:Permissions@acm.org}{Permissions@acm.org}. \\
% \emph{CHI '16},  May 07--12, 2016, San Jose, CA, USA \\
% ACM xxx-x-xxxx-xxxx-x/xx/xx\ldots \$15.00 \\
% DOI: \url{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
% }
\toappear{}

% Arabic page numbers for submission.  Remove this line to eliminate
% page numbers for the camera ready copy
% \pagenumbering{arabic}

% Load basic packages
\usepackage{balance}       % to better equalize the last page
\usepackage{graphics}      % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
\usepackage{txfonts}
\usepackage{mathptmx}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}

% Some optional stuff you might like/need.
\usepackage{microtype}        % Improved Tracking and Kerning
% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
\usepackage{ccicons}          % Cite your images correctly!
% \usepackage[utf8]{inputenc} % for a UTF8 editor only

% If you want to use todo notes, marginpars etc. during creation of
% your draft document, you have to enable the "chi_draft" option for
% the document class. To do this, change the very first line to:
% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
% by using the "\todo{...}"  command. Make sure to disable the draft
% option again before submitting your final document.
\usepackage{todonotes}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{Visualizing NBA teams across multiple seasons}
\def\plainauthor[#1]{#1}
\def\plainauthors[#1]#2#3{#1, #2, #3}
\def\emptyauthor{}
\def\plainkeywords{Authors' choice; of terms; separated; by
  semicolons; include commas, within terms only; required.}
\def\plaingeneralterms{Documentation, Standardization}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
  pdfauthor={\plainauthors[Yannick Boesmans]{Yannick Laevaert}{Yves Langeraert}},
%  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  pdfdisplaydoctitle=true, % For Accessibility
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{\plaintitle}

\numberofauthors{3}
\author{%
  \alignauthor{\plainauthor[Yannick Boesmans]\\
    \affaddr{KU Leuven}\\
    \email{yannick.boesmans}\\
    \email{@student.kuleuven.be}}\\
  \alignauthor{\plainauthor[Yannick Laevaert]\\
    \affaddr{KU Leuven}\\
    \email{yannick.laevaert}\\
    \email{@student.kuleuven.be}}\\
  \alignauthor{\plainauthor[Yves Langeraert]\\
    \affaddr{KU Leuven}\\
    \email{yves.langeraert}\\
    \email{@student.kuleuven.be}}\\
}

\maketitle

\begin{abstract}
    This paper details the conceptualization and implementation of an
    interactive visualization of data about NBA seasons including teams, game
    results and player statistics. The d3.js framework was used to develop this
    visualization.
    %TODO add results
\end{abstract}

%TODO keywords
\category{H.5.m.}{Information Interfaces and Presentation
  (e.g. HCI)}{Miscellaneous} \category{See
  \url{http://acm.org/about/class/1998/} for the full list of ACM
  classifiers. This section is required.}{}{}

\keywords{\plainkeywords}

\section{Introduction}

This paper describes the visualization of NBA data made by the ``The Tufters''
team for the course ``Information Visualization''. In section~\ref{sec:goal}
we describe the goal of the visualization and the target audience. In
section~\ref{sec:data} we describe the data used, including its origins,
advantages and limitations. In section~\ref{sec:literature} we give an overview
of related literature and web resources, including related visualizations.
Related visualizations include both visualizations of NBA or other sports data,
as well as visualizations tackling a problem we encountered during the
development of our visualization. In section~\ref{sec:visualization} we describe
the visualization itself. We give an overview of the different stages of
development of the visualization, as well as the major design decisions made
during the development process. Section~\ref{sec:discussion} discusses potential
improvements of the final visualization and lessons learned from the project. We
conclude in section~\ref{sec:conclusion} by giving a short overview.

\section{Goal and Target Audience}\ref{sec:goal}
The visualization's goal is allowing exploration of NBA data by lay persons.
More specifically, the visualization does not try to offer premade explanations
for phenomenons visible in the NBA data. By providing easy and intuitive access
to the data, users can draw their own conclusions. 
Secondly, the visualization's target audience are lay persons. This means people
not professionally active in the field of basketball. Specifically, fans of NBA
are the core of our target audience.

\section{Data}\ref{sec:data}
The data visualized is a subset of the data available on basketball statistics
site basketball-reference\cite{basketball-reference}. A wide range of data is
available on this site. In our visualization we only use data from 1971 onwards.
The data we use includes league standings and playoff rankings for each team,
the team's roster and individual player statistics for each year, including the
PER~(Player Efficiency Rating).

\section{Related Work}\ref{sec:literature}
\section{Visualization}\ref{sec:visualization}
\section{Discussion}\ref{sec:discussion}
\section{Conclusion}\ref{sec:conclusion}

% Balancing columns in a ref list is a bit of a pain because you
% either use a hack like flushend or balance, or manually insert
% a column break.  http://www.tex.ac.uk/cgi-bin/texfaq2html?label=balance
% multicols doesn't work because we're already in two-column mode,
% and flushend isn't awesome, so I choose balance.  See this
% for more info: http://cs.brown.edu/system/software/latex/doc/balance.pdf
%
% Note that in a perfect world balance wants to be in the first
% column of the last page.
%
% If balance doesn't work for you, you can remove that and
% hard-code a column break into the bbl file right before you
% submit:
%
% http://stackoverflow.com/questions/2149854/how-to-manually-equalize-columns-
% in-an-ieee-paper-if-using-bibtex
%
% Or, just remove \balance and give up on balancing the last page.
%
\balance{}

% REFERENCES FORMAT
% References must be the same font size as other body text.
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{sample}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: